Another Look at Regularized Discriminant Analysis for High-Dimensional Classification

### 2013-04-21 reports/paper-rda.Rnw
Removed bolding of vectors and matrices to match Biometrika's requirements.

\newcommand{\xbar}{\bar{\bm x}}
\newcommand{\tr}{\text{tr}}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}

@@

\subsection{Preliminaries}
In discriminant analysis we wish to assign correctly an unlabeled
$p$-dimensional observation vector $\bm x$ to one of $K$ unique, known classes
(or populations) by constructing a classifier from $n$ training observations
that can accurately predict the class membership of $\bm x$. Let $\bm x_i =
(x_{i1}, \ldots, x_{ip}) \in \mathbb{R}_{p \times 1}$ be the $i$th observation
$(i = 1, \ldots, N)$ with true, unique membership $y_i \in \{\omega_1, \ldots,
\omega_K\}$, where $\mathbb{R}_{a \times b}$ denotes the matrix space of all $a
\times b$ matrices over the real field $\mathbb{R}$. We assume that $(\bm x_i,
y_i)$ is a realization from a mixture distribution $p(\bm x) = \sum_{k=1}^K
p(\bm x | \omega_k) p(\omega_k)$, where $p(\bm x | \omega_k)$ is the probability
density function of the $k$th class and $p(\omega_k)$ is the prior probability
of class membership of the $k$th class. We further assume that each of
$p(\omega_k)$ is equal $(k = 1, \ldots, K)$.

Quadratic discriminant analysis is the optimal Bayesian decision rule with
respect to a $0-1$ loss function when $p(\bm x | \omega_k)$ is the probability
density function of the multivariate normal distribution with known mean vector
$\bm\mu_k \in \mathbb{R}_{p \times 1}$ and known covariance matrix $\bm\Sigma_k
\in \mathbb{R}_{p \times p}^{>}$, $k = 1, 2, \ldots, K$, where $\mathbb{R}_{p
  \times p}^{>}$ denotes the cone of real $p \times p$ positive-definite
matrices. Because $\bm \mu_k$ and $\bm\Sigma_k$ are typically unknown, we assign
an unlabeled observation $\bm x$ to class $\omega_k$ with the sample quadratic
discriminant analysis classifier
\begin{align}
	D_{QDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\widehat{\bm\Sigma}_k^{-1}(\bm x - \xbar_k)  + \log |\widehat{\bm\Sigma}_k|, \label{eq:qda}
\end{align}
where $\xbar_k$ and $\widehat{\bm\Sigma}_k$ are the maximum likelihood
estimators for $\bm \mu_k$ and $\bm \Sigma_k$, respectively. If we assume
further that $\bm\Sigma_k = \bm\Sigma$ $(k = 1, \ldots, K)$, then the pooled
sample covariance matrix $\widehat{\bm\Sigma}$ is substituted for $\widehat{\bm
  \Sigma}_k$ in \eqref{eq:qda}, where
\begin{align}
	\widehat{\bm\Sigma} = N^{-1} \sum_{k=1}^K n_k \widehat{\bm\Sigma}_k \label{eq:pooled-cov}
\end{align}
is the maximum likelihood estimator for $\bm \Sigma$. In this case
\eqref{eq:qda} reduces to the sample linear discriminant analysis classifier,
and the log-determinant is omitted because it is constant across all classes.

Throughout the document we also require the following additional notation. Let
$\bm A \oplus \bm B$ denote the direct sum of $\bm A \in \mathbb{R}_{r \times
  r}$ and $\bm B \in \mathbb{R}_{s \times s}$ \citep[Chapter
  1]{Lutkepohl:1996uz}. Let $\bm A^+$ and $\mathcal{N}(\bm A) = \{\bm z \in
\mathbb{R}_{p \times 1} : \bm A \bm z = \bm 0\}$ denote the Moore-Penrose
pseudoinverse and null space of $\bm A \in \mathbb{R}_{m \times p}$,
respectively. Let $V^{\perp} = \{\bm y \in \mathbb{R}_{p \times 1} : \bm y'\bm v
= 0\ \forall \bm v \in V \}$ denote the orthogonal complement of a vector space
$V \subset \mathbb{R}_{p \times 1}$. Let $\bm I_m$ be the $m \times m$ identity
matrix, and let $\bm 0_{m \times p}$ be the $m \times p$ matrix of zeros, such
that $\bm 0_m$ is understood to denote $\bm 0_{m \times m}$. Lastly, for $c \in
\mathbb{R}$, let $c^+ = 1/c$ if $c \ne 0$ and $0$ otherwise.

@@

The smallest eigenvalues and the directions associated with their eigenvectors
highly influence \eqref{eq:qda}. In fact, the eigenvalues of $\widehat{\bm
  \Sigma}_k$ are well-known to be biased if $p \ge n_k$ such that the smallest
eigenvalues are underestimated \citep{Seber:2004uh}, Moreover, if $p > n_k$,
then rank$(\widehat{\bm \Sigma}_k) \le n_k$, which implies that at least $p -
n_k$ eigenvalues of $\widehat{\bm \Sigma}_k$ are 0. Furthermore, although more
feature information is available to discriminate among the $K$ classes, if $p >
n_k$, \eqref{eq:qda} is incalculable because $\widehat{\bm \Sigma}_k^{-1}$ does
not exist.

Several regularization methods, such as the methods considered by
\citet*{Guo:2007te}, \cite{Mkhadri:1995jp}, and \citet*{Xu:2009fl}, have been
proposed in the literature to adjust the eigenvalues of $\widehat{\bm \Sigma}_k$
so that \eqref{eq:qda} is calculable and has reduced variability in the standard
bias-variance tradeoff. A common form of the covariance-matrix regularization
applies a shrinkage factor $\gamma > 0$, so that
\begin{align}
	\widehat{\bm \Sigma}_k(\gamma) =  \widehat{\bm \Sigma}_k + \gamma \bm I_p, \label{eq:ridge-estimator}
\end{align}
similar to a method employed in ridge regression \citep{Hoerl:1970cd}. Equation
\eqref{eq:ridge-estimator} effectively \emph{shrinks} the sample covariance
matrix $\widehat{\bm\Sigma}_k$ towards $\bm I_p$, thereby increasing the
eigenvalues of $\widehat{\bm\Sigma}_k$ by $\gamma$. Specifically, the zero
eigenvalues are replaced with $\gamma$, so that \eqref{eq:ridge-estimator} is
positive definite. For additional covariance-matrix regularization methods, see
\cite{Ramey:2013ji}, \cite{Xu:2009fl}, and \cite{Ye:2009gd}.

@@

\cite{Friedman:1989tm} has proposed regularized discriminant analysis by
incorporating a weighted average of the sample covariance matrix $\widehat{\bm
  \Sigma}_k$ and the pooled sample covariance matrix $\widehat{\bm\Sigma}$ to
estimate $\bm \Sigma_k$ with
\begin{align}
  \widehat{\bm\Sigma}_k(\lambda) = n_k^{-1}(\lambda) \bm S_k(\lambda) \quad (k =
  1, \ldots, K),\label{eq:sig-lambda}
\end{align}
where $\lambda \in [0, 1]$, $n_k(\lambda) = (1 - \lambda) n_k + \lambda N$, $\bm
S_k = n_k \widehat{\bm\Sigma}_k$, $\bm S = \sum_{k=1}^K \bm S_k$, and $\bm
S_k(\lambda) = (1 - \lambda) \bm S_k + \lambda \bm S$. We can interpret
\eqref{eq:sig-lambda} as a covariance-matrix estimator that borrows from
$\widehat{\bm\Sigma}$ in \eqref{eq:pooled-cov} to better estimate $\bm
\Sigma_k$. If $\lambda = 0$, \eqref{eq:sig-lambda} is equal to
$\widehat{\bm\Sigma}_k$ used in \eqref{eq:qda}. Alternatively, if $\lambda = 1$,
\eqref{eq:sig-lambda} reduces to \eqref{eq:pooled-cov}.

To further improve the estimation of $\bm \Sigma_k$ and to stabilize the inverse
of \eqref{eq:sig-lambda}, \cite{Friedman:1989tm} has proposed the biased
covariance-matrix estimator
\begin{align}
	\widehat{\bm\Sigma}_k(\lambda, \gamma) = (1 - \gamma) \widehat{\bm\Sigma}_k(\lambda) + \gamma \frac{\tr\{\widehat{\bm\Sigma}_k(\lambda)\}}{p} \bm I_p,\label{eq:sig-rda}
\end{align}
where $\gamma \in [0, 1]$ is a regularization parameter that controls the
shrinkage of \eqref{eq:sig-rda} towards $\bm I_p$ and is weighted by the average
of the eigenvalues of \eqref{eq:sig-lambda}. Thus, the \emph{pooling} parameter
$\lambda$ controls the amount borrowed from $\widehat{\bm\Sigma}$ to estimate
$\bm \Sigma_k$, and the \emph{shrinkage} parameter $\gamma$ determines the
amount of eigenvalue shrinkage applied. Finally, by substituting
\eqref{eq:sig-rda} into \eqref{eq:qda}, we have the regularized discriminant
analysis classifier
\begin{align}
	D_{RDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\widehat{\bm\Sigma}_k(\lambda, \gamma)^{-1}(\bm x - \xbar_k)  + \log |\widehat{\bm\Sigma}_k(\lambda, \gamma)|. \label{eq:rda}
\end{align}
If $\gamma$ is sufficiently small, \eqref{eq:sig-rda} can be singular or
near-singular. For this case \cite{Friedman:1989tm} has suggested replacing the
zero eigenvalues of \eqref{eq:sig-rda} with a small positive constant that
permits a numerically stable matrix inversion.

@@

Here, we define our adaptation of \citeauthor{Friedman:1989tm}'s classifier by
first simplifying the formulation of $\widehat{\bm \Sigma}_k(\lambda)$ in
\eqref{eq:sig-lambda} and demonstrating its clear interpretation as a linear
combination of the crossproducts of the training observations centered by their
respective class sample means. We define the convex combination
\begin{align}
  \widehat{\bm\Sigma}_k(\lambda) = (1 - \lambda) \widehat{\bm\Sigma}_k + \lambda
  \widehat{\bm\Sigma}, \quad (k = 1, \ldots, K)\label{eq:sig-lambda-alternative}
\end{align}
where $\lambda \in [0, 1]$ is the \emph{pooling} parameter. Thus, we avoid the
convex combination of $n_k$ and $N$ as defined by \cite{Friedman:1989tm}. By
rewriting \eqref{eq:sig-lambda-alternative} in terms of the observations $\bm
x_i$ $(i = 1, \ldots, N)$, each centered by its class sample mean, we attain a
clear interpretation of $\widehat{\bm\Sigma}_k(\lambda)$. That is,
\begin{align}
	\widehat{\bm\Sigma}_k(\lambda) &= \widehat{\bm\Sigma}_k + \lambda (\widehat{\bm\Sigma} - \widehat{\bm\Sigma}_k) \nonumber\\
%	&= \widehat{\bm\Sigma}_k + \lambda \left(\sum_{k' = 1}^K \frac{n_k}{N} \widehat{\bm\Sigma}_{k'}  - \widehat{\bm\Sigma}_k \right) \nonumber\\
	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right) \widehat{\bm\Sigma}_k +  \frac{\lambda}{N} \sum_{\substack{k' = 1\\k' \ne k}} n_{k'} \widehat{\bm\Sigma}_{k'} \nonumber \\
%	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right)\sum_{i=1}^N I(y_i = k) \bm x_i \bm x_i' +  \frac{\lambda}{N} \sum_{i=1}^N I(y_i \ne k) \bm x_i \bm x_i' \nonumber \\
	&= \sum_{i=1}^N c_{ik}(\lambda) \bm x_i \bm x_i',\label{eq:sig-lambda-alternative2}
\end{align}
where $c_{ik}(\lambda) = \lambda N^{-1} + (1 - \lambda)n_k^{-1}I(y_i = k)$.
From \eqref{eq:sig-lambda-alternative2}, $\lambda$ weights the contribution of
each of the $N$ observations to estimate $\bm \Sigma_k$. Hence, for $0 < \lambda
\le 1$, we estimate $\bm \Sigma_k$ with $N$ rather than $n_k$
observations. Below, we will show that the pooling operation is advantageous in
increasing the rank of each $\widehat{\bm\Sigma}_k(\lambda)$ from
rank$(\widehat{\bm\Sigma}_k)$ to rank$(\widehat{\bm\Sigma})$ for $0 < \lambda
\le 1$. Notice that if $\lambda = 0$, then the observations from the remaining
$K - 1$ classes do not contribute to the estimation of $\bm \Sigma_k$,
corresponding to $\widehat{\bm \Sigma}_k$. Furthermore, if $\lambda = 1$, the
weights in \eqref{eq:sig-lambda-alternative2} reduce to $1/N$, corresponding to
$\widehat{\bm\Sigma}$. For brevity, when $\lambda = 1$, we define $\bm X =
[\sqrt{c_{1k}(1)} \bm x_1', \ldots, \sqrt{c_{Nk}(1)} \bm x_N']'$ such that
$\widehat{\bm \Sigma} = \bm X' \bm X$. Similarly, for $\lambda = 0$, we define
$\bm X_k = [\sqrt{c_{1k}(0)} \bm x_1', \ldots, \sqrt{c_{Nk}(0)} \bm x_N']'$ such
that $\widehat{\bm \Sigma}_k = \bm X_k' \bm X_k$. In Figure
\ref{fig:hdrda-contours} we plot the contours of five multivariate normal
populations for $\lambda = 0$ with unequal covariance matrices. As $\lambda$
approaches 1, the contours become more similar, resulting in identical contours
for $\lambda = 1$.

@@

As we have discussed above, several eigenvalue adjustment methods have been
proposed to increase eigenvalues (approximately) equal to 0. Here, we define the
eigenvalue adjustment of \eqref{eq:sig-lambda-alternative} as
\begin{align}
	\tilde{\bm \Sigma}_k = \alpha_k \widehat{\bm \Sigma}_k(\lambda) + \gamma \bm I_p,\label{eq:hdrda-cov}
\end{align}
where $\alpha_k \ge 0$ and $\gamma \ge 0$ is an eigenvalue-shrinkage
constant. The choice of $\alpha_k$ allows for a flexible formulation of
covariance-matrix estimators. For instance, if $\alpha_k = 1$, $k = 1, \ldots,
K$, then \eqref{eq:hdrda-cov} resembles \eqref{eq:ridge-estimator}. Similarly,
if $\alpha_k = 1 - \gamma$, then \eqref{eq:hdrda-cov} has a form similar to
\eqref{eq:sig-rda}. Now, substituting \eqref{eq:hdrda-cov} into \eqref{eq:qda},
we define the high-dimensional regularized discriminant analysis classifier
\begin{align}
	D_{HDRDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\tilde{\bm\Sigma}_k^{+}(\bm x - \xbar_k)  + \log |\tilde{\bm\Sigma}_k|. \label{eq:hdrda}
\end{align}
For $\gamma > 0$, $\tilde{\bm\Sigma}_k$ is nonsingular such that
$\tilde{\bm\Sigma}_k^{-1}$ can be substituted for $\tilde{\bm\Sigma}_k^{+}$ in
\eqref{eq:hdrda}. If $\gamma = 0$, we explicitly set $|\tilde{\bm\Sigma}_k|$
equal to the product of the positive eigenvalues of $\tilde{\bm\Sigma}_k$.

@@

Next, we demonstrate that our proposed classifier in \eqref{eq:hdrda} lends
itself to a more efficient calculation and a rapid model selection of the
parameters $\lambda$ and $\gamma$\footnote{TODO: Briefly mention the updating
  trick employed in the R package}. We decompose \eqref{eq:hdrda} into a sum of
two components, where the first summand consists of matrix operations applied to
low-dimensional matrices and the second summand corresponds to the null space of
$\widehat{\bm \Sigma}$ in \eqref{eq:pooled-cov}. We show that the matrix
operations performed on the null space of $\widehat{\bm \Sigma}$ yield constant
quadratic forms across all classes and can be omitted. For $p \gg N$, the
constant component involves determinants and inverses of high-dimensional
matrices, and by ignoring these calculations, a substantial reduction in
computational costs is achieved relative to Friedman's classifier in
\eqref{eq:rda}\footnote{TODO: Add a brief comparison of the computations of
  \emph{RDA} and \emph{HDRDA} in terms of Big Oh notation}. Furthermore, a
byproduct is that adjustments to the associated eigenvalues have no effect on
\eqref{eq:hdrda}. Lastly, we utilize the singular value decomposition to
calculate efficiently the eigenvalue decomposition of $\widehat{\bm \Sigma}$,
further reducing the computational costs of our proposed classifier.

First, we require the following relationship regarding the null spaces of
$\widehat{\bm \Sigma}_k(\lambda)$, $\widehat{\bm \Sigma}$, and $\widehat{\bm
  \Sigma}_k$.

>>>>

\newcommand{\xbar}{\bar{x}}
\newcommand{\tr}{\text{tr}}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}

@@

\subsection{Preliminaries}
In discriminant analysis we wish to assign correctly an unlabeled
$p$-dimensional observation vector $x$ to one of $K$ unique, known classes
(or populations) by constructing a classifier from $N$ training
observations. Let $x_i = (x_{i1}, \ldots, x_{ip}) \in \mathbb{R}_{p \times
  1}$ be the $i$th observation $(i = 1, \ldots, N)$ with true, unique membership
$y_i \in \{\omega_1, \ldots, \omega_K\}$, where $\mathbb{R}_{a \times b}$
denotes the matrix space of all $a \times b$ matrices over the real field
$\mathbb{R}$. We assume that $(x_i, y_i)$ is a realization from a mixture
distribution $p(x) = \sum_{k=1}^K p(x | \omega_k) p(\omega_k)$, where
$p(x | \omega_k)$ is the probability density function of the $k$th class and
$p(\omega_k)$ is the prior probability of class membership of the $k$th
class. We further assume that each of $p(\omega_k)$ is equal $(k = 1, \ldots,
K)$.

Quadratic discriminant analysis is the optimal Bayesian decision rule with
respect to a $0-1$ loss function when $p(x | \omega_k)$ is the probability
density function of the multivariate normal distribution with known mean vector
$\mu_k \in \mathbb{R}_{p \times 1}$ and known covariance matrix $\Sigma_k
\in \mathbb{R}_{p \times p}^{>}$, $k = 1, 2, \ldots, K$, where $\mathbb{R}_{p
  \times p}^{>}$ denotes the cone of real $p \times p$ positive-definite
matrices. Because $\mu_k$ and $\Sigma_k$ are typically unknown, we assign
an unlabeled observation $x$ to class $\omega_k$ with the sample quadratic
discriminant analysis classifier
\begin{align}
	D_{QDA}(x) = \argmin_{k}  (x - \xbar_k)'\widehat{\Sigma}_k^{-1}(x - \xbar_k)  + \log |\widehat{\Sigma}_k|, \label{eq:qda}
\end{align}
where $\xbar_k$ and $\widehat{\Sigma}_k$ are the maximum likelihood
estimators for $\mu_k$ and $\Sigma_k$, respectively. If we assume
further that $\Sigma_k = \Sigma$ $(k = 1, \ldots, K)$, then the pooled
sample covariance matrix $\widehat{\Sigma}$ is substituted for $\widehat{
  \Sigma}_k$ in \eqref{eq:qda}, where
\begin{align}
	\widehat{\Sigma} = N^{-1} \sum_{k=1}^K n_k \widehat{\Sigma}_k \label{eq:pooled-cov}
\end{align}
is the maximum likelihood estimator for $\Sigma$. In this case \eqref{eq:qda}
reduces to the sample linear discriminant analysis classifier, and the
log-determinant is omitted because it is constant across all classes.

Throughout the document we also require the following additional notation. Let
$I_m$ be the $m \times m$ identity matrix, and let $0_{m \times p}$ be
the $m \times p$ matrix of zeros, such that $0_m$ is understood to denote
$0_{m \times m}$. Lastly, for $c \in \mathbb{R}$, let $c^+ = 1/c$ if $c \ne
0$ and $0$ otherwise. Let $A \oplus B$ denote the direct sum of $A \in
\mathbb{R}_{r \times r}$ and $B \in \mathbb{R}_{s \times s}$ \citep[Chapter
  1]{Lutkepohl:1996uz}. Let $A^+$ and $\mathcal{N}(A) = \{z \in \mathbb{R}_{p
  \times 1} : A z = 0_{m \times 1}\}$ denote the Moore-Penrose pseudoinverse and null space
of $A \in \mathbb{R}_{m \times p}$, respectively. Let $V^{\perp} = \{y
\in \mathbb{R}_{p \times 1} : y'v = 0\ \forall v \in V \}$ denote
the orthogonal complement of a vector space $V \subset \mathbb{R}_{p \times 1}$.

@@

The smallest eigenvalues and the directions associated with their eigenvectors
highly influence \eqref{eq:qda}. In fact, the eigenvalues of $\widehat{
  \Sigma}_k$ are well-known to be biased if $p \ge n_k$ such that the smallest
eigenvalues are underestimated \citep{Seber:2004uh}, Moreover, if $p > n_k$,
then rank$(\widehat{\Sigma}_k) \le n_k$, which implies that at least $p -
n_k$ eigenvalues of $\widehat{\Sigma}_k$ are 0. Furthermore, although more
feature information is available to discriminate among the $K$ classes, if $p >
n_k$, \eqref{eq:qda} is incalculable because $\widehat{\Sigma}_k^{-1}$ does
not exist.

Several regularization methods, such as the methods considered by
\citet*{Guo:2007te}, \cite{Mkhadri:1995jp}, and \citet*{Xu:2009fl}, have been
proposed in the literature to adjust the eigenvalues of $\widehat{\Sigma}_k$
so that \eqref{eq:qda} is calculable and has reduced variability in the standard
bias-variance tradeoff. A common form of the covariance-matrix regularization
applies a shrinkage factor $\gamma > 0$, so that
\begin{align}
	\widehat{\Sigma}_k(\gamma) =  \widehat{\Sigma}_k + \gamma I_p, \label{eq:ridge-estimator}
\end{align}
similar to a method employed in ridge regression \citep{Hoerl:1970cd}. Equation
\eqref{eq:ridge-estimator} effectively \emph{shrinks} the sample covariance
matrix $\widehat{\Sigma}_k$ towards $I_p$, thereby increasing the
eigenvalues of $\widehat{\Sigma}_k$ by $\gamma$. Specifically, the zero
eigenvalues are replaced with $\gamma$, so that \eqref{eq:ridge-estimator} is
positive definite. For additional covariance-matrix regularization methods, see
\cite{Ramey:2013ji}, \cite{Xu:2009fl}, and \cite{Ye:2009gd}.

@@

\cite{Friedman:1989tm} has proposed regularized discriminant analysis by
incorporating a weighted average of the sample covariance matrix $\widehat{
  \Sigma}_k$ and the pooled sample covariance matrix $\widehat{\Sigma}$ to
estimate $\Sigma_k$ with
\begin{align}
  \widehat{\Sigma}_k(\lambda) = n_k^{-1}(\lambda) S_k(\lambda) \quad (k =
  1, \ldots, K),\label{eq:sig-lambda}
\end{align}
where $\lambda \in [0, 1]$, $n_k(\lambda) = (1 - \lambda) n_k + \lambda N$, $
S_k = n_k \widehat{\Sigma}_k$, $S = \sum_{k=1}^K S_k$, and $
S_k(\lambda) = (1 - \lambda) S_k + \lambda S$. We can interpret
\eqref{eq:sig-lambda} as a covariance-matrix estimator that borrows from
$\widehat{\Sigma}$ in \eqref{eq:pooled-cov} to better estimate $
\Sigma_k$. If $\lambda = 0$, \eqref{eq:sig-lambda} is equal to
$\widehat{\Sigma}_k$ used in \eqref{eq:qda}. Alternatively, if $\lambda = 1$,
\eqref{eq:sig-lambda} reduces to \eqref{eq:pooled-cov}.

To further improve the estimation of $\Sigma_k$ and to stabilize the inverse
of \eqref{eq:sig-lambda}, \cite{Friedman:1989tm} has proposed the biased
covariance-matrix estimator
\begin{align}
	\widehat{\Sigma}_k(\lambda, \gamma) = (1 - \gamma) \widehat{\Sigma}_k(\lambda) + \gamma \frac{\tr\{\widehat{\Sigma}_k(\lambda)\}}{p} I_p,\label{eq:sig-rda}
\end{align}
where $\gamma \in [0, 1]$ is a regularization parameter that controls the
shrinkage of \eqref{eq:sig-rda} towards $I_p$ and is weighted by the average
of the eigenvalues of \eqref{eq:sig-lambda}. Thus, the \emph{pooling} parameter
$\lambda$ controls the amount borrowed from $\widehat{\Sigma}$ to estimate
$\Sigma_k$, and the \emph{shrinkage} parameter $\gamma$ determines the
amount of eigenvalue shrinkage applied. Finally, by substituting
\eqref{eq:sig-rda} into \eqref{eq:qda}, we have the regularized discriminant
analysis classifier
\begin{align}
	D_{RDA}(x) = \argmin_{k}  (x - \xbar_k)'\widehat{\Sigma}_k(\lambda, \gamma)^{-1}(x - \xbar_k)  + \log |\widehat{\Sigma}_k(\lambda, \gamma)|. \label{eq:rda}
\end{align}
If $\gamma$ is sufficiently small, \eqref{eq:sig-rda} can be singular or
near-singular. For this case \cite{Friedman:1989tm} has suggested replacing the
zero eigenvalues of \eqref{eq:sig-rda} with a small positive constant that
permits a numerically stable matrix inversion.

@@

Here, we define our adaptation of \citeauthor{Friedman:1989tm}'s classifier by
first simplifying the formulation of $\widehat{\Sigma}_k(\lambda)$ in
\eqref{eq:sig-lambda} and demonstrating its clear interpretation as a linear
combination of the crossproducts of the training observations centered by their
respective class sample means. We define the convex combination
\begin{align}
  \widehat{\Sigma}_k(\lambda) = (1 - \lambda) \widehat{\Sigma}_k + \lambda
  \widehat{\Sigma}, \quad (k = 1, \ldots, K)\label{eq:sig-lambda-alternative}
\end{align}
where $\lambda \in [0, 1]$ is the \emph{pooling} parameter. Thus, we avoid the
convex combination of $n_k$ and $N$ as defined by \cite{Friedman:1989tm}. By
rewriting \eqref{eq:sig-lambda-alternative} in terms of the observations $
x_i$ $(i = 1, \ldots, N)$, each centered by its class sample mean, we attain a
clear interpretation of $\widehat{\Sigma}_k(\lambda)$. That is,
\begin{align}
	\widehat{\Sigma}_k(\lambda) &= \widehat{\Sigma}_k + \lambda (\widehat{\Sigma} - \widehat{\Sigma}_k) \nonumber\\
%	&= \widehat{\Sigma}_k + \lambda \left(\sum_{k' = 1}^K \frac{n_k}{N} \widehat{\Sigma}_{k'}  - \widehat{\Sigma}_k \right) \nonumber\\
	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right) \widehat{\Sigma}_k +  \frac{\lambda}{N} \sum_{\substack{k' = 1\\k' \ne k}} n_{k'} \widehat{\Sigma}_{k'} \nonumber \\
%	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right)\sum_{i=1}^N I(y_i = k) x_i x_i' +  \frac{\lambda}{N} \sum_{i=1}^N I(y_i \ne k) x_i x_i' \nonumber \\
	&= \sum_{i=1}^N c_{ik}(\lambda) x_i x_i',\label{eq:sig-lambda-alternative2}
\end{align}
where $c_{ik}(\lambda) = \lambda N^{-1} + (1 - \lambda)n_k^{-1}I(y_i = k)$.
From \eqref{eq:sig-lambda-alternative2}, $\lambda$ weights the contribution of
each of the $N$ observations to estimate $\Sigma_k$. Hence, for $0 < \lambda
\le 1$, we estimate $\Sigma_k$ with $N$ rather than $n_k$
observations. Below, we will show that the pooling operation is advantageous in
increasing the rank of each $\widehat{\Sigma}_k(\lambda)$ from
rank$(\widehat{\Sigma}_k)$ to rank$(\widehat{\Sigma})$ for $0 < \lambda
\le 1$. Notice that if $\lambda = 0$, then the observations from the remaining
$K - 1$ classes do not contribute to the estimation of $\Sigma_k$,
corresponding to $\widehat{\Sigma}_k$. Furthermore, if $\lambda = 1$, the
weights in \eqref{eq:sig-lambda-alternative2} reduce to $1/N$, corresponding to
$\widehat{\Sigma}$. For brevity, when $\lambda = 1$, we define $X =
[\sqrt{c_{1k}(1)} x_1', \ldots, \sqrt{c_{Nk}(1)} x_N']'$ such that
$\widehat{\Sigma} = X' X$. Similarly, for $\lambda = 0$, we define
$X_k = [\sqrt{c_{1k}(0)} x_1', \ldots, \sqrt{c_{Nk}(0)} x_N']'$ such
that $\widehat{\Sigma}_k = X_k' X_k$. In Figure
\ref{fig:hdrda-contours} we plot the contours of five multivariate normal
populations for $\lambda = 0$ with unequal covariance matrices. As $\lambda$
approaches 1, the contours become more similar, resulting in identical contours
for $\lambda = 1$.

@@

As we have discussed above, several eigenvalue adjustment methods have been
proposed to increase eigenvalues (approximately) equal to 0. Here, we define the
eigenvalue adjustment of \eqref{eq:sig-lambda-alternative} as
\begin{align}
	\tilde{\Sigma}_k = \alpha_k \widehat{\Sigma}_k(\lambda) + \gamma I_p,\label{eq:hdrda-cov}
\end{align}
where $\alpha_k \ge 0$ and $\gamma \ge 0$ is an eigenvalue-shrinkage
constant. The choice of $\alpha_k$ allows for a flexible formulation of
covariance-matrix estimators. For instance, if $\alpha_k = 1$, $k = 1, \ldots,
K$, then \eqref{eq:hdrda-cov} resembles \eqref{eq:ridge-estimator}. Similarly,
if $\alpha_k = 1 - \gamma$, then \eqref{eq:hdrda-cov} has a form similar to
\eqref{eq:sig-rda}. Now, substituting \eqref{eq:hdrda-cov} into \eqref{eq:qda},
we define the high-dimensional regularized discriminant analysis classifier
\begin{align}
	D_{HDRDA}(x) = \argmin_{k}  (x - \xbar_k)'\tilde{\Sigma}_k^{+}(x - \xbar_k)  + \log |\tilde{\Sigma}_k|. \label{eq:hdrda}
\end{align}
For $\gamma > 0$, $\tilde{\Sigma}_k$ is nonsingular such that
$\tilde{\Sigma}_k^{-1}$ can be substituted for $\tilde{\Sigma}_k^{+}$ in
\eqref{eq:hdrda}. If $\gamma = 0$, we explicitly set $|\tilde{\Sigma}_k|$
equal to the product of the positive eigenvalues of $\tilde{\Sigma}_k$.

@@

Next, we demonstrate that our proposed classifier in \eqref{eq:hdrda} lends
itself to a more efficient calculation and a rapid model selection of the
parameters $\lambda$ and $\gamma$\footnote{TODO: Briefly mention the updating
  trick employed in the R package}. We decompose \eqref{eq:hdrda} into a sum of
two components, where the first summand consists of matrix operations applied to
low-dimensional matrices and the second summand corresponds to the null space of
$\widehat{\Sigma}$ in \eqref{eq:pooled-cov}. We show that the matrix
operations performed on the null space of $\widehat{\Sigma}$ yield constant
quadratic forms across all classes and can be omitted. For $p \gg N$, the
constant component involves determinants and inverses of high-dimensional
matrices, and by ignoring these calculations, a substantial reduction in
computational costs is achieved relative to Friedman's classifier in
\eqref{eq:rda}\footnote{TODO: Add a brief comparison of the computations of
  \emph{RDA} and \emph{HDRDA} in terms of Big Oh notation}. Furthermore, a
byproduct is that adjustments to the associated eigenvalues have no effect on
\eqref{eq:hdrda}. Lastly, we utilize the singular value decomposition to
calculate efficiently the eigenvalue decomposition of $\widehat{\Sigma}$,
further reducing the computational costs of our proposed classifier.

First, we require the following relationship regarding the null spaces of
$\widehat{\Sigma}_k(\lambda)$, $\widehat{\Sigma}$, and $\widehat{
  \Sigma}_k$.

\begin{lemma}\label{lemma:null-spaces}
Let $\widehat{\Sigma}_k$ and $\widehat{\Sigma}$ be the maximum
likelihood estimators of $\Sigma_k$ and $\Sigma$, respectively. Let
$\widehat{\Sigma}_k(\lambda)$ be defined as in
\eqref{eq:sig-lambda-alternative}. Then, $\mathcal{N}\{\widehat{\Sigma}_k(\lambda)\} \subset \mathcal{N}(\widehat{
  \Sigma}) \subset \mathcal{N}(\widehat{\Sigma}_k)$ $(k = 1, \ldots, K)$.
\end{lemma}
\begin{proof}
Let $z \in \mathcal{N}\{\widehat{\Sigma}_k(\lambda)\}$ for some $k = 1,
\ldots, K$. Hence, $0 = z' \widehat{\Sigma}_k(\lambda) z = (1 -
\lambda) z' \widehat{\Sigma}_k z + \lambda z' \widehat{
  \Sigma} z$. Because $\widehat{\Sigma}_k, \widehat{\Sigma}\in
\mathbb{R}_{p \times p}^{\ge}$, we have $z \in \mathcal{N}(\widehat{
  \Sigma})$ and $z \in \mathcal{N}(\widehat{\Sigma}_k)$. In particular,
we have that $\mathcal{N}\{\widehat{\Sigma}_k(\lambda)\} \subset
\mathcal{N}(\widehat{\Sigma})$. Now, suppose $z \in
\mathcal{N}(\widehat{\Sigma})$. Similarly, we have that $0 = z'
\widehat{\Sigma} z = N^{-1} \sum_{k = 1}^K n_k z' \widehat{
  \Sigma}_k z$, which implies that $z \in \mathcal{N}(\widehat{
  \Sigma}_k)$ because $\widehat{\Sigma}_k \in \mathbb{R}_{p \times
  p}^{\ge}$. Therefore, $\mathcal{N}(\widehat{\Sigma}) \subset
\mathcal{N}(\widehat{\Sigma}_k)$.
\end{proof}

In Lemma \ref{lemma:rda-tilde-Sigma_k} below, we derive an alternative
expression for $\tilde{\Sigma}_k$ in terms of the matrix of eigenvectors of
$\widehat{\Sigma}$. Let $\widehat{\Sigma} = U D U'$ be the
eigendecomposition of $\widehat{\Sigma}$ such that $D \in \mathbb{R}_{p
  \times p}^{\ge}$ is the diagonal matrix of eigenvalues of $\widehat{\Sigma}$
with $D = D_q \oplus 0_{p-q}$, $D_q \in \mathbb{R}_{q \times
  q}^{>}$ is the diagonal matrix consisting of the positive eigenvalues of
$\widehat{\Sigma}$, the columns of $U \in \mathbb{R}_{p \times p}$ are
the corresponding orthonormal eigenvectors of $\widehat{\Sigma}$, and
rank$(\widehat{\Sigma}) = q$. Then, we partition $U = (U_1, 
U_2)$ such that $U_1 \in \mathbb{R}_{p \times q}$ and $U_2 \in
\mathbb{R}_{p \times (p - q)}$.
